### Question 0

*  Draw a concepts diagram that uses all the following Github terminology 
   * Include any missing keywords that will simplify the concepts diagrams 


`Push`, `Repository`, `Clone`, `Pull`, `Pull Request`, `Branch`, `Merging`, `Github Client`, `README file`, `Private or Public`

```{r}
# Add your photo here
https://drive.google.com/file/d/1A1x-xynrE4IM46lqo1tGgDFvsFRSUcsl/view?usp=drive_link 
```

#### Question 1

`matrix(c(1,2,3,4,5,6) , nrow = 3)`

* Running the expression produces the following matrix

|  |  |
|---|---|
| 1 | 4 |
| 2 | 5 |
| 3 | 6 |

* How can you modify the call to `matrix()` to produce the following matrix instead?

|   | col_1 |  col_2 |
| --- |---|---|
| row_1 | 1 | 2 |
| row_2 | 3 | 4 |
| row_3 | 5 | 6 |

* Note that you need to name the columns (col_1 and col_2) and name the rows (row_1, row_2, row_3)

Hint: Use the `?` symbol to invoke the matrix documentation

```{r}
# Add your code here
practical_2_matrix<- matrix(c(1,3,5,2,4,6) , nrow = 3)
colnames(practical_2_matrix)<- c("col_1", "col_2")
rownames(practical_2_matrix)<- c("row_1", "row_2", "row_3")
practical_2_matrix
```


#### Question 2

* Load then sort the airquality data frame on its `Temp` and `Solar.R` columns in reverse order (largest to smallest values)
  * The function to sort a data frame is called order
* Display only the first 15 lines of your table

```{r}
data(airquality)
head(airquality)
#create temp and solar vectors in order to perform the order function (becuase it has to be a vector)
temp_vector<- airquality[ , c("Temp")]
solar_vector<- airquality[ , c("Solar.R")]
#reverse sort temp and solar vectors 
temp_sorted<- temp_vector[order(-temp_vector)]
solar_sorted<- solar_vector[order(-solar_vector)]
#bind sorted temp and solar vectors into a data frame
sorted_df<- data.frame(temp_sorted, solar_sorted)
#display only first 15 lines of new data frame
head(sorted_df,15)

#retry this following example from class (different from above)
data(airquality)
head(airquality)

df_temp<- airquality[order(-airquality$Temp), ]
df_solar<- airquality[order(-airquality$Solar.R), ]
decreasing_df<- data.frame(df_temp, df_solar)
head(decreasing_df, 15)
```

#### Question 3

* Sort the airquality data frame on its `Temp` in decreasing order and `Solar.R` in increasing order
* Display only the first 15 lines of your table

```{r}
#write your answer here
data(airquality)
head(airquality)
#create temp and solar vectors in order to perform the order function (becuase it has to be a vector)
temp_vector<- airquality[ , c("Temp")]
solar_vector<- airquality[ , c("Solar.R")]
#sort temp and solar vectors 
temp_sorted<- temp_vector[order(-temp_vector)]
solar_sorted<- solar_vector[order(solar_vector)]
#bind sorted temp and solar vectors into a data frame
sorted_df<- data.frame(temp_sorted, solar_sorted)
#display only first 15 lines of new data frame
head(sorted_df,15)

#retry this one with the example from class (different than above)
data(airquality)
head(airquality)

new_df_temp<- airquality[order(-airquality$Temp), ]
new_df_solar<- airquality[order(airquality$Solar.R), ]
new_df<- data.frame(new_df_temp, new_df_solar)
new_df
head(new_df,15)

```

#### Question 4

* There are various ways to select a subset of observations from a data frame.

* Consult your [R Reference Card](https://cran.r-project.org/doc/contrib/Baggott-refcard-v2.pdf), see `Data Selection and Manipulation` section.
  * What operations can you use to select all observations where the temperature is 72. Give at least two different answers to this question


```{r}
# Write your answer here
data(airquality)
head(airquality)

#1st way
temp_sorted<- temp_sorted==72 
new_temp<- temp_sorted
filtered_new_temp<- temp_sorted[new_temp]

#2nd way
airquality[is.element(airquality$Temp, c(72)), ]

```

#### Question 6

* You may have noticed when working with the `airqulity` data that some values show as `NA`
 * `NA` stands for not available, or missing values.
* A major part of data wrangling consists of cleaning missing values by either:
  * Dropping the lines that have missing values
   * Sometimes we can drop the column with missing values if the column is made of predominantly missing values
  * Imputing the missing values, which uses educated guesses (or more complex algorithms) to fill the missing values
  
* Find and remove all rows that are missing values for the `Solar.R` or `Ozone` variables
* Save the cleaned data to a new data frame called airquality_no_na
  * How many lines have been removed?

```{r}
# First attempt on my own
#data(airquality)
#head(airquality)
#create vectors for columns 
#solar_vector<- airquality[ , c("Solar.R")]
#ozone_vector<- airquality[ , c("Ozone")]
#remove NAs from columns 
#solar_no_na<- na.omit(solar_vector)
#ozone_no_na<- na.omit(ozone_vector)
#airquality_no_na<- data.frame(solar_no_na, ozone_no_na) #this doesn't work because they dont have a matching number of rows--ask about this#

#solar now has 146 and ozone has 116, so solar is missing 7 and ozone is missing 37  

#final attempt following example from class--avoided the issue I had above: 
data(airquality)
head(airquality)

is.na(airquality$Solar.R & airquality$Ozone)
airquality_no_na<- airquality[!is.na(airquality$Solar.R & airquality$Ozone), ]
airquality_no_na

#solar now has 146 and ozone has 116, so solar is missing 7 and ozone is missing 37 

```

#### Question 7
* Let's use a different strategy and impute the missing value.
  * replace the missing values for Solar.R using that month's average.
  * Example:
    * The missing value for line 6 should be replaced with the average for month 5.
    * The missing value for line 97 should be replaced with the average for month 8.

```{r}
# Write your answer here
data(airquality)
head(airquality)

#create average value for months with NA values (May and August)
may_obs<- subset(airquality, Month == 5, "Solar.R")
may_obs_no_na<- na.omit(may_obs)
may_average<- mean(may_obs_no_na$Solar.R)

aug_obs<- subset(airquality, Month == 8, "Solar.R")
aug_obs_no_na<- na.omit(aug_obs)
aug_average<- mean(aug_obs_no_na$Solar.R)

#replace NAs for May and August with average values 
may_obs[is.na(may_obs)] <- may_average
aug_obs[is.na(aug_obs)] <- aug_average
```

